---
title: "Final Report for Intro to Data Science"
author: "Griffin Sun"
output: html_notebook
---

# Introduction
<p> Hello, I am Griffin Sun and this is my Final project for Intro to Data science. The data set I am using has collated the top 50 songs on Spotify in 2019 and characterizes the songs by certain attributes such as </p>
  * Dance-ability
  * Energy
  * Loudness
  
<p> We all know the four chords that pervade modern music but I want to see if attributes like a songs dance-ability or energy have an effect on its popularity. </p> 

# Preparation

lock and load the packages
```{r}
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("dplyr")
library(tidyverse)
library(ggplot2)
library(dplyr)
```

Load the data
```{r}
library(readr)
dat <- read_csv("top50.csv")
```

# Analysis

## Averages and Range


```{r}
#prepare the data to chart
dance<- "#ff825c"
energy<- "#88a44e"
loud<- "#1e9e9e"
BP<- "#c7993b"
val<- "#a1a141"

dat%>%
  ggplot()+
  geom_boxplot(aes(x="danceability",y=Danceability,color=dance,))+
  geom_boxplot(aes(x="Energy",y=Energy,color=energy))+
  geom_boxplot(aes(x="BPM",y=Beats.Per.Minute,color=BP))+
  geom_boxplot(aes(x="Valence",y=Valence.,color=val))+
  labs(y="value", legend.position="none", )

```

<p> This chart tells us that over the past that at an attribute level, the ingredients for a summer hit haven't changed.</p>

```{r}
dat%>%
  ggplot()+
  geom_boxplot(aes(x="Loudness",y=Loudness,color=loud))
```
```{r}

```

